manuscript_figures

### 2024-06-26 scripts/manuscript_figures.Rmd
calcualating pb and hpi ratios

# Load libraries 
```{r}
library(plyr)
library(kableExtra)
library(gt)
library(dplyr)
library(tigris)
library(MASS)
library(htmlwidgets)
library(car)
library(ggiraphExtra)
library(MuMIn)
library(lmerTest)
library(tidyverse)
library(lme4)
library(broom.mixed)
library(ggmap)
library(lubridate)
library(cowplot)
library(ggplot2)
library(PerformanceAnalytics)
library(reshape2)
library(viridis)
library(GGally)
library(corrplot)
library(patchwork)
library(broom)
library(knitr)
library(ggspatial)
library(rslp)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggspatial)

@@

@@

# Make line plot for temperature time series by site per month
```{r}
# Calculate average temperature per month and site
avg_temp <- temperature_data %>%
  mutate(MonthAbbrev = month.abb[Month]) %>%  # Add month abbreviations
  group_by(MonthAbbrev, Site) %>%
  summarize(avg_temp = mean(Temp))

@@

# Plotting average temperature by site as lines and points per month
temperature_plot <- ggplot(avg_temp, aes(x = MonthAbbrev, y = avg_temp, color = Site, group = Site)) +
  geom_line(size = 1.5) +  # Lines for average temperature by site
  geom_point(size = 3) +  # Points for mean values by site
  labs(x = "Month", y = expression("Average Temperature (°C)"), color = "Site") +
  scale_color_manual(values = c("Fort Wetherill" = "grey", "King's Beach" = "black")) +  # Custom colors
  theme_bw()+
    theme(
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size = 18),
    axis.text.x = element_text(size = 18),
    axis.text.y = element_text(size = 18),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 18));temperature_plot

@@

Production to biomass ratios (P:B) for each fish species
```{r}

>>>>

# Load libraries 
```{r}
library(plyr)
library(kableExtra)
library(flextable)
library(gt)
library(dplyr)
library(tigris)
library(MASS)
library(htmlwidgets)
library(car)
library(ggiraphExtra)
library(MuMIn)
library(lmerTest)
library(tidyverse)
library(lme4)
library(broom.mixed)
library(ggmap)
library(lubridate)
library(cowplot)
library(ggplot2)
library(PerformanceAnalytics)
library(reshape2)
library(viridis)
library(GGally)
library(corrplot)
library(patchwork)
library(broom)
library(knitr)
library(ggspatial)
library(rslp)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggspatial)

@@

# Identify the outliers
outliers <- c(27,7)

# remove outliers
data_clean <- aggregated_data_cover[-outliers, ]

# rerun model with log and without outliers
lmm_model <- lmer(log(mean_fish_biomass + 1.1) ~ log(mean_kelp_cover + 1.1) + YEAR + (1 | SITE), data = data_clean)
qqPlot(residuals(lmm_model)) #improved to normal
summary(lmm_model)

# Extract coefficients and their standard errors
model_summary <- tidy(lmm_model)

# Create a data frame with coefficients, standard errors, and p-values
df <- data.frame(
  Coefficient = model_summary$term,
  Estimate = model_summary$estimate,
  Std.Error = model_summary$std.error,
  P.Value = model_summary$p.value
)

# Create flextable object
flex <- flextable(df)

# Save flextable as PNG
png_file <- "output/model.result.tables/rugosity.not.included/summary_fish_kelp_cover_year_fixed.png"
save_as_image(flex, png_file)

# rerun model with log and without outliers
lmm_model <- lmer(log(mean_fish_biomass + 1.1) ~ log(mean_kelp_cover + 1.1) * YEAR + (1 | SITE), data = data_clean)
qqPlot(residuals(lmm_model)) #improved to normal
summary(lmm_model)

# Extract coefficients and their standard errors
model_summary <- tidy(lmm_model)

# Create a data frame with coefficients, standard errors, and p-values
df <- data.frame(
  Coefficient = model_summary$term,
  Estimate = model_summary$estimate,
  Std.Error = model_summary$std.error,
  P.Value = model_summary$p.value
)

# Create flextable object
flex <- flextable(df)

# Save flextable as PNG
png_file <- "output/model.result.tables/rugosity.not.included/summary_fish_kelp_cover_year_interaction.png"
save_as_image(flex, png_file)

# rerun model with log and without outliers
lmm_model <- lmer(log(mean_fish_biomass + 1.1) ~ log(mean_kelp_cover + 1.1) + SITE + (1 | SITE/YEAR), data = data_clean)
qqPlot(residuals(lmm_model)) #improved to normal
summary(lmm_model)

# Extract coefficients and their standard errors
model_summary <- tidy(lmm_model)

# Create a data frame with coefficients, standard errors, and p-values
df <- data.frame(
  Coefficient = model_summary$term,
  Estimate = model_summary$estimate,
  Std.Error = model_summary$std.error,
  P.Value = model_summary$p.value
)

# Create flextable object
flex <- flextable(df)

# Save flextable as PNG
png_file <- "output/model.result.tables/rugosity.not.included/summary_fish_kelp_cover.png"
save_as_image(flex, png_file)

@@

# Make line plot for temperature time series by site per month
```{r}
# Calculate mean and standard error of temperature by month and site
avg_temp <- temperature_data %>%
  mutate(MonthAbbrev = month.abb[Month]) %>%  # Add month abbreviations
  group_by(MonthAbbrev, Site) %>%
  summarize(
    avg_temp = mean(Temp),
    std_error = sd(Temp) / sqrt(n()))  # Calculate standard error

@@

# Plotting average temperature by site as lines and points per month with error bars
temperature_plot <- ggplot(avg_temp, aes(x = MonthAbbrev, y = avg_temp, color = Site, group = Site)) +
  geom_line(size = 1.5) +  # Lines for average temperature by site
  geom_point(size = 3) +  # Points for mean values by site
  geom_errorbar(aes(ymin = avg_temp - std_error, ymax = avg_temp + std_error), width = 0.2) +  # Error bars for SEM
  labs(x = "Month", y = expression("Average Temperature (°C)"), color = "Site") +
  scale_color_manual(values = c("Fort Wetherill" = "grey", "King's Beach" = "black")) +  # Custom colors
  theme_bw() +
  theme(
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size = 18),
    axis.text.x = element_text(size = 18),
    axis.text.y = element_text(size = 18),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 18));temperature_plot

@@

Production to biomass ratios (P:B) for each fish species
```{r}
# load fish data from 2016
PB <- read.csv("data/fish.data.clean.calc.csv")

# remove habitat and control, month, day, date, not needed
PB <- PB[ -c(1:2, 4:7) ]

# Calculate fish biomass 'W' using the length-weight relationship
PB$biomass <- with(PB, coef.a * (actual_size)^coef.b)

# Divide biomass by 80 m to get g/m2
#calculate biomass/m2
PB$biomass.g.m2 <- PB$biomass / 80

# Constants from the P/B ratio equation
coefficient <- 2.64
exponent <- -0.35

# Calculate P/B ratio
PB$P_B_ratio <- coefficient * PB$biomass.g.m2^exponent

# Calculate Habitat Productivity Index (HPI)
PB$HPI <- PB$biomass.g.m2 * PB$P_B_ratio

# Aggregate by Transect, Site, Year and calculate mean HPI
hpi_summary <- PB %>%
  group_by(TRANSECT, SITE, YEAR) %>%
  summarise(mean_HPI = mean(HPI))

ggplot(hpi_summary, aes(x = paste(TRANSECT, SITE, sep = " - "), y = mean_HPI, fill = as.factor(YEAR))) +
  geom_bar(stat = "identity", position = "dodge", color = "black") +
  labs(x = "Transect - Site", y = "Mean Habitat Productivity Index (HPI)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
